A Graph Theory Based Approach to MDAO Problem Formulation

### 2013-03-15 graph_syntax.tex
more updates

\subsection{MDAO Graph Primatives}
There are three node types:  
\begin{description}
\item[variable node] represents passing of data.
\item[model node] represents the handling of data.
\item[driver node] represents constrol structures capable of managing iteration
\end{description}
There are two edge types: 
\begin{description}
\item[fixed edge] A fixed edge may not be removed.
\item[free edge] A free edge may be removed.
\end{description}

@@

The MDO problem formulation concepts represented by nodes are given in 
Table \ref{t:node representation}, and the concepts represented by edges are 
given in Table \ref{t:edge representation}.
% \begin{itemize}
% \item An objective or constraint is indicated by a variable node with no outgoing edges and with the incoming edges being directed from only other variable nodes.
% \item A global input is a variable node with no edges directed in and with at least two edges directed out to different variable nodes.
% \item A design variable is represented by a variable node with no edges directed in and on fixed edge directed out.
% \end{itemize}
\begin{table}[h!]
 \begin{center}
  \caption{Problem formulation concept represented by a node}
  \label{t:node representation}
  \begin{tabular}{ccc} \hline 
edges directed in & edges directed out & node representation \\ \hline
only free edges & none & objective or constraint\\
none & at least two free edges & global input \\
none & one fixed edge & design variable \\ \hline
  \end{tabular}
 \end{center}
 \vspace{-15pt}
\end{table}
\begin{table}[h!]
 \begin{center}
  \caption{Problem formulation concept represented by an edge}
  \label{t:edge representation}
  \begin{tabular}{ccc} \hline 
from node & to node & edge representation \\ \hline
variable & variable & connection/passing of a variable\\
variable & model & local input \\
model & variable & local output \\
model & model & function call \\ \hline
  \end{tabular}
 \end{center}
 \vspace{-15pt}
\end{table}

Finally, an analysis code is represented by an \emph{analysis block}, which is a collection of model nodes, variable nodes, and fixed edges with a specific structure. This structure is derived from the need to represent each input and output individually while representing the actual analysis with a single node or edge. The variable nodes each represent a single input or output of the analysis block. Each of these nodes is connected to a single model node via a fixed edge, which represents the gathering of the inputs to begin the analysis or the disseminating of outputs after the analysis. Finally, free edges are used to connect the variable nodes of the analysis block to other variables nodes outside of the analysis block. In this way, the analysis block graph is a fundamental building block of the MDAO problem data flow.
\begin{figure}[htb!]
	\begin{center}
	\includegraphics[width=4in]{images/analysis_block}
	\end{center}
	\vspace{-10pt}
\caption{Example analysis block. The each node type and edge type is labeled in italics and annotated parenthetically.}
\label{f:analysis block}
\end{figure}
%`Connection' edges connect the local outputs from one analysis block to input nodes.

>>>>

\subsection{MDAO Graph Primatives}
There are three node types:  
\begin{description}
\item[variable node] represents passing of data.
\item[model node] represents the handling of data.
\item[driver node] represents constrol structures capable of managing iteration
\end{description}
There are two edge types: 
\begin{description}
\item[fixed edge] A fixed edge may never be removed from the graph.
\item[free edge] A free edge may be removed removed from the graph.
\end{description}

@@

Analysis tools take in a number input variables, then perform some work calculate 
the values for their respective outputs. In an MDAO graph, this process is 
represented by an group of nodes and edges called an \emph{analysis block}, 
shown in Fig. \ref{f:analysis block}. Within an analysis block each variable 
node represents a single input or output and is connected 
to a single model node via a fixed edge. Note that in Fig. \ref{f:analysis block}, 
the analysis block contains two model nodes, with a single edge connecting them. 
This part of the graph represents the necessary calculations to map given inputs 
to proper output values.

\begin{figure}[htb!]
    \begin{center}
    \includegraphics[width=4in]{images/analysis_block}
    \end{center}
    \vspace{-10pt}
\caption{Example analysis block. The each node type and edge type is labeled in italics and annotated parenthetically.}
\label{f:analysis block}
\end{figure}

Finally, free edges are used to connect the variable nodes of the analysis block 
to other variables nodes outside of the analysis block. In this way, the 
analysis block graph is a fundamental building block of the MDAO problem data flow.

The MDO problem formulation concepts represented by nodes are given in 
Table \ref{t:node representation}, and the concepts represented by edges are 
given in Table \ref{t:edge representation}.
% \begin{itemize}
% \item An objective or constraint is indicated by a variable node with no outgoing edges and with the incoming edges being directed from only other variable nodes.
% \item A global input is a variable node with no edges directed in and with at least two edges directed out to different variable nodes.
% \item A design variable is represented by a variable node with no edges directed in and on fixed edge directed out.
% \end{itemize}
\begin{table}[h!]
 \begin{center}
  \caption{Problem formulation concept represented by a node}
  \label{t:node representation}
  \begin{tabular}{ccc} \hline 
edges directed in & edges directed out & node representation \\ \hline
only free edges & none & objective or constraint\\
none & at least two free edges & global input \\
none & one fixed edge & design variable \\ \hline
  \end{tabular}
 \end{center}
 \vspace{-15pt}
\end{table}
\begin{table}[h!]
 \begin{center}
  \caption{Problem formulation concept represented by an edge}
  \label{t:edge representation}
  \begin{tabular}{ccc} \hline 
from node & to node & edge representation \\ \hline
variable & variable & connection/passing of a variable\\
variable & model & local input \\
model & variable & local output \\
model & model & function call \\ \hline
  \end{tabular}
 \end{center}
 \vspace{-15pt}
\end{table}
%`Connection' edges connect the local outputs from one analysis block to input nodes.
